---
title: "Statistical Distributions and Shrinkage Priors"
author: "Duong Trinh"
date: "`r Sys.Date()`"
output:
  html_document: 
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
  pdf_document: 
    number_sections: true
    extra_dependencies: ["mathtools","bbm"]
runtime: shiny
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
options(knitr.duplicate.label = "allow")
```

<!--- For HTML Only --->
`r if (knitr:::is_html_output()) '
$\\newcommand{\\mathbbm}[1]{\\mathbb{#1}}$
'`

```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```


# Statistical Distributions

## Normal Distribution

The normal distribution has density:

$$
f(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}} exp\left(-\frac{(x-\mu)^2}{2\sigma^2}\right)
$$

```{r, warning=FALSE}
help("Normal")
```

Plotting the standard normal distribution

```{r}
# library(shiny)
# library(ggplot2)
# library(tidyverse)
# library(miniUI)
# 
# 
# myGadgetFunc <- function() {
#   
#   ui <- miniPage(
#     gadgetTitleBar("My Gadget"),
#     miniContentPanel(
#       # Define layout, inputs, outputs
#     sliderInput('Mean','Mean', min = - 3, max = 3, value = 0, step = 1),
#     sliderInput('Sigma','Sigma', min = 1, max = 5, value = 1, step = 1),
#     plotOutput('plotDensity')
#     )
#   )
#   
#   serve <- function(input, output, session) {
#     # Define reactive expressions, outputs, etc.
#     selectedData <- reactive({
#     data_frame(x = c(input$Mean - 4*input$Sigma, input$Mean + 4*input$Sigma))
#   })
#   
#   output$plotDensity <- renderPlot({
#     ggplot(selectedData(), aes(x=x)) +
#       geom_vline(xintercept = input$Mean, linetype = "dashed", alpha = 0.4) +
#       stat_function(fun = dnorm, args = c(input$Mean,input$Sigma),
#                     color = "#84CA72", size = 1) +
#       ggtitle(paste0("Normal Distribution N (", input$Mean," , ", input$Sigma^2, ")")) +
#       xlab("x") +
#       ylab("dnorm(x)") +
#       coord_cartesian(ylim = c(0,0.5))
#   })
#     # When the Done button is clicked, return a value
#     observeEvent(input$done, {
#       returnValue <- ...
#       stopApp(returnValue)
#     })
#     
#   }
#   
#   runGadget(ui,serve)
# }
```

```{r}
library(shiny)
library(ggplot2)
library(tidyverse)

ui <- fluidPage(
  column(4,
    wellPanel(
    sliderInput('Mu','Mu', min = - 3, max = 3, value = 0, step = 1),
    sliderInput('Sigma','Sigma', min = 1, max = 5, value = 1, step = 1)
    )
  ),
  column(8,
    plotOutput('plotDensity')
  )
)

server <- function(input,output) {
  
  selectedData <- reactive({
    data_frame(x = c(input$Mu - 4*input$Sigma, input$Mu + 4*input$Sigma))
  })
  
  output$plotDensity <- renderPlot({
    ggplot(selectedData(), aes(x=x)) +
      geom_vline(xintercept = input$Mu, linetype = "dashed", alpha = 0.4) +
      stat_function(fun = dnorm, args = c(input$Mu,input$Sigma),
                    color = "#84CA72", size = 1) +
      labs(caption = paste0("Normal Distribution N (", input$Mu," , ", input$Sigma^2, ")")) +
      xlab("x") +
      ylab("dnorm(x)") +
      coord_cartesian(ylim = c(0,0.5)) +
      theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      ) 
      
  })
  
}

shinyApp(ui = ui, server = server)
```


```{r, warning=FALSE, fig.align='center', results='hide', fig.show='hide'}
mu <- 0
sigma <- 1

p <- data_frame(x = c(mu - 4*sigma, mu + 4*sigma)) %>% 
  ggplot(aes(x=x))

p + geom_vline(xintercept = mu, linetype = "dashed", alpha = 0.4) +
    stat_function(fun = dnorm, args = c(mu,sigma),
                  color = "#84CA72", size = 1) +
    ggtitle(paste0("Normal Distribution N (", mu," , ", sigma^2, ")")) +
    xlab("x") +
    ylab("dnorm(x)") 
```


```{r, results='hide', fig.show='hide'}
ggplot(data.frame(x = -5:5), aes(x)) + 
  stat_function(fun = dnorm, args = c(0,1), geom = 'area', xlim = c(1,5), fill = 4) + 
  stat_function(fun = dnorm, args = c(0,1)) +
  labs(x = "", y = "")+
  ggtitle("Normal distribution")
```

```{r}
library("Rcpp")

set.seed(123)
rnorm(1,0,1)

set.seed(123)
evalCpp("R::rnorm(0,1)")

set.seed(123)
rnorm(3,0,1)

set.seed(123)
evalCpp("Rcpp::rnorm(3,0,1)")
```


```{r}
ndraws <- 1e6

ggplot(data_frame(samples = rnorm(ndraws)), aes(x = samples)) + 
  geom_density(fill = "lightgreen") +
  stat_function(fun = function(x) dnorm(x), color = "red")
```


## *t* distribution
The *t* distribution with `df` = $\nu$ degrees of freedom has density

$$
f(x) = 
$$

```{r, warning=FALSE}
help("TDist")
```

Plotting the Central *t* distribution

```{r}
library(shiny)
library(ggplot2)
library(tidyverse)

ui <- fluidPage(
  column(4,
    wellPanel(
    sliderInput('df','df', min = 1, max = 12, value = 6, step = 1),
    sliderInput('ncp','ncp', min = 1, max = 5, value = 1, step = 1)
    )
  ),
  column(8,
    plotOutput('plotDensity')
  )
)

server <- function(input,output) {
  
  mean <- 0
  sd <- reactive({sqrt(input$df/(input$df-2))})
  selectedData <- reactive({
    data_frame(x = c(mean - 4*sd(), mean + 4*sd()))
  })
  
  output$plotDensity <- renderPlot({
    ggplot(selectedData(), aes(x=x)) +
      geom_vline(xintercept = mean, linetype = "dashed", alpha = 0.4) +
      stat_function(fun = dt, args = c(input$df),
                    color = "#84CA72", size = 1) +
      labs(caption = paste0("t Distribution (df = ", input$df,")")) +
      xlab("x") +
      ylab(paste0("dt(x, df = ", input$df,")")) +
      coord_cartesian(ylim = c(0,0.5)) +
      theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      ) 
      
  })
  
}

shinyApp(ui = ui, server = server)
```


```{r, warning=FALSE, fig.align='center', results='hide', fig.show='hide'}
df <- 6
ncp <- NA # non-centrality parameter

mean <- 0
sd <- sqrt(df/(df-2))

p <- data_frame(x = c(mean - 4*sd, mu + 4*sd)) %>% 
  ggplot(aes(x=x))

p + geom_vline(xintercept = mean, linetype = "dashed", alpha = 0.4) +
    stat_function(fun = dt, args = c(df),
                  color = "#84CA72", size = 1) +
    labs(caption = paste0("t Distribution (df = ", df,")")) +
    xlab("x") +
    ylab(paste0("dt(x, df = ", df,")")) +
    coord_cartesian(ylim = c(0,0.5)) +
    theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      ) 
```

## Gamma Distribution

The Gamma distribution with parameters `shape` = $\alpha$ and `scale` = $\sigma$ has density

$$
f(x \mid \alpha, \sigma) = \frac{x^{\alpha - 1} e^{(-x/\sigma)}}{\sigma^\alpha \Gamma(\alpha)} 
$$
for $x \geq 0, \alpha > 0$ and $\sigma > 0$.

The mean and variance are

$$
E(X) = \alpha \sigma; \quad Var(X) = \alpha \sigma^2
$$

```{r, warning=FALSE}
help("GammaDist")
```


Plotting the Gamma distribution

```{r}
library(shiny)
library(ggplot2)
library(tidyverse)

ui <- fluidPage(
  column(4,
    wellPanel(
    sliderInput('alpha','alpha (shape)', min = 1, max = 12, value = 1, step = 1),
    sliderInput('sigma','sigma (scale)', min = 0, max = 3, value = 0.5, step = 0.1)
    )
  ),
  column(8,
    plotOutput('plotDensity')
  )
)

server <- function(input,output) {
  
  mean <- reactive({input$alpha*input$sigma})
  sd <- reactive({sqrt(input$alpha*input$sigma^2)})

  selectedData <- reactive({
    data_frame(x = c(0, qgamma(p = 0.9999, shape = input$alpha, scale = input$sigma)))
  })
  
  output$plotDensity <- renderPlot({
    ggplot(selectedData(), aes(x=x)) +
      geom_vline(xintercept = mean(), linetype = "dashed", alpha = 0.4) +
      stat_function(fun = dgamma, args = c(input$alpha,input$sigma),
                    color = "#84CA72", size = 1) +
      labs(caption = paste0("Gamma (alpha = ", input$alpha, ", sigma = ", input$sigma,") Distribution")) +
      xlab("x") +
      ylab(paste0("dgamma(x, alpha = ", input$alpha, ", sigma = ", input$sigma,")")) +
      coord_cartesian(ylim = c(0,0.5)) +
      theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      ) 
      
  })
  
}

shinyApp(ui = ui, server = server)
```


```{r, warning=FALSE, fig.align='center', results='hide', fig.show='hide'}
# dgamma(x, shape, rate = 1, scale = 1/rate, log = FALSE)
alpha <- 1 # shape
sigma <- 0.5 # scale
rate <- 1/sigma # 1/scale  


mean <- alpha*sigma
sd <- sqrt(alpha*sigma^2)

p <- data_frame(x = c(0, qgamma(p = 0.9999, shape = alpha, scale = sigma))) %>% 
  ggplot(aes(x=x))

p + geom_vline(xintercept = mean, linetype = "dashed", alpha = 0.4) +
    stat_function(fun = dgamma, args = c(shape = alpha, scale = sigma),
                  color = "#84CA72", size = 1) +
    ggtitle(paste0("Gamma (alpha = ", alpha, ", sigma = ", sigma,") Distribution")) +
    xlab("x") +
    ylab(paste0("dgamma(x, alpha = ", alpha, ", sigma = ", sigma,")"))
```


```{r}
library("Rcpp")

set.seed(123)
rgamma(3,1,1)

set.seed(123)
Rcpp::evalCpp("randg(3, arma::distr_param(1,1))", depends = "RcppArmadillo")

vecgamma <- cppFunction(depends = "RcppArmadillo",
                  '
  arma::vec C() {
    return randg(3, arma::distr_param(1,1));
  }
'
)
set.seed(123)
vecgamma()
```

## Inverse-Gamma Distribution
The Inverse-Gamma distribution with parameters `shape` = $\alpha$ and `scale` = $\beta$ has density

$$
f(x \mid \alpha, \beta) = \frac{\beta^\alpha x^{-\alpha-1} exp(-\beta/x)}{\Gamma(\alpha)} 
$$
for $x \geq 0, \alpha > 0$ and $\beta > 0$.

The mean and variance are

$$
E(X) = \frac{\beta}{\alpha-1} \quad (\alpha>1) ; \quad Var(X) = \frac{\beta^2}{(\alpha-1)^2(\alpha-2)} \quad (\alpha>2)
$$

Plotting the Inverse-Gamma distribution

```{r}
library(shiny)
library(ggplot2)
library(tidyverse)

ui <- fluidPage(
  column(4,
    wellPanel(
    sliderInput('alpha','alpha (shape)', min = 1, max = 12, value = 1, step = 1),
    sliderInput('beta','beta (scale)', min = 0, max = 3, value = 0.5, step = 0.5)
    )
  ),
  column(8,
    plotOutput('plotDensity')
  )
)

server <- function(input,output) {
  
  mode <- reactive({input$beta/(input$alpha+1)})

  selectedData <- reactive({
    data_frame(x = c(0, qinvgamma(p = 0.9, shape = input$alpha, scale = input$beta)))
  })
  
  output$plotDensity <- renderPlot({
    ggplot(selectedData(), aes(x=x)) +
      geom_vline(xintercept = mode(), linetype = "dashed", alpha = 0.4, color = "blue") +
      stat_function(fun = dinvgamma, args = c(input$alpha,input$beta),
                    color = "#84CA72", size = 1) +
      labs(caption = paste0("Inverse Gamma (alpha = ", input$alpha, ", beta = ", input$beta,") Distribution")) +
      xlab("x") +
      ylab(paste0("dinvgamma(x, alpha = ", input$alpha, ", beta = ", input$beta,")")) +
      coord_cartesian(xlim = c(0,3)) +
      coord_cartesian(ylim = c(0,5)) +
      theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      ) 
      
  })
  
}

shinyApp(ui = ui, server = server)
```

```{r, warning=FALSE}
alpha <- 1 # shape
beta <- 1 # scale

# install.packages("invgamma")
# library(LaplacesDemon)
# library(MCMCpack)
library(invgamma)
q9999 <- qinvgamma(p = 0.9999, shape = alpha, scale = beta)
p <- data_frame(x = c(0, q9999)) %>% 
  ggplot(aes(x=x))

p + geom_vline(xintercept = mean, linetype = "dashed", alpha = 0.4) +
    stat_function(fun = dinvgamma, args = c(shape = alpha, scale = beta),
                  color = "#84CA72", size = 1) +
    ggtitle(paste0("Inverse Gamma (alpha = ", alpha, ", beta = ", beta,") Distribution")) +
    xlab("x") +
    ylab(paste0("dinvgamma (x, alpha = ", alpha, ", beta = ", beta,")"))
```

```{r}
ndraws <- 1e6

# rinvgamma(ndraws, shape = alpha, scale = beta)
alpha <- 1 # shape
beta <- 1 # scale

g1 <- rgamma(ndraws, shape = alpha, scale = 1/beta)
g2 <- 1/g1

data_frame(g1=g1,g2=g2) %>% summarise_each(funs(mean, min,  max, sd))

plotg1 <- ggplot(data_frame(g1), aes(x=g1)) + 
  # geom_density(fill = "lightgreen") +
  stat_function(fun = function(x) dgamma(x, shape = alpha, scale = 1/beta), color = "red")

plotg2 <- ggplot(data_frame(g2), aes(x=g2)) + 
  # geom_density(fill = "green") +
  stat_function(fun = function(x) dinvgamma(x, shape = alpha, scale = beta), color = "red") +
  xlim(0,10)

plotg3 <- ggplot(data_frame(g3 = rgamma(ndraws, shape = alpha, scale = 1/beta)), aes(x=g3)) + 
  # geom_density(fill = "lightgreen") +
  stat_function(fun = function(x) dgamma(x, shape = alpha, scale = 1/beta), color = "red")

plotg4 <- ggplot(data_frame(g4 = rinvgamma(ndraws, shape = alpha, scale = beta)), aes(x=g4)) + 
  # geom_density(fill = "green") +
  stat_function(fun = function(x) dinvgamma(x, shape = alpha, scale = beta), color = "red") +
  xlim(0,10)


library(gridExtra)
grid.arrange(plotg1,plotg2, ncol = 2)
grid.arrange(plotg3,plotg4, ncol = 2)
```


## Exponential Distribution
The exponential distribution with parameter `rate` = $\lambda$ has density

$$
f(x \mid \lambda) = \lambda e^{-\lambda x}
$$

for $x \geq 0$.

The mean and variance are

$$
E(X) = \frac{1}{\lambda}; \quad Var(X) = \frac{1}{\lambda^2}
$$

Plotting the Exponential distribution

```{r}
library(shiny)
library(ggplot2)
library(tidyverse)

ui <- fluidPage(
  column(4,
    wellPanel(
    sliderInput('lambda','lambda (rate)', min = 0.5, max = 2, value = 1, step = 0.5),
    )
  ),
  column(8,
    plotOutput('plotDensity')
  )
)

server <- function(input,output) {
  
  mean <- reactive({1/input$lambda})
  sd <- reactive({1/input$lambda})

  selectedData <- reactive({
    data_frame(x = c(0, qexp(p = 0.9999, rate = input$lambda)))
  })
  
  output$plotDensity <- renderPlot({
    ggplot(selectedData(), aes(x=x)) +
      geom_vline(xintercept = mean(), linetype = "dashed", alpha = 0.4) +
      stat_function(fun = dexp, args = c(input$lambda),
                    color = "#84CA72", size = 1) +
      labs(caption = paste0("Exponential (lambda = ", input$lambda,") Distribution")) +
      xlab("x") +
      ylab(paste0("dexp(x, lambda = ", input$lambda,")")) +
      coord_cartesian(ylim = c(0,2)) +
      theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      ) 
      
  })
  
}

shinyApp(ui = ui, server = server)
```

```{r, warning=FALSE, fig.align='center', results='hide', fig.show='hide'}
# dexp(x, rate = 1, log = FALSE)
lambda <- 1 # rate

mean <- 1/lambda
sd <- 1/lambda

p <- data_frame(x = c(0, qexp(p = 0.9999, rate = lambda))) %>% 
  ggplot(aes(x=x))

p + geom_vline(xintercept = mean, linetype = "dashed", alpha = 0.4) +
    stat_function(fun = dexp, args = c(rate = lambda),
                  color = "#84CA72", size = 1) +
    ggtitle(paste0("Exponential (lambda = ", lambda,") Distribution")) +
    xlab("x") +
    ylab(paste0("dexp(x, lambda = ", lambda,")")) +
    coord_cartesian(ylim = c(0,1.5))
```

## Laplace Distribution

The Laplace distribution with parameters `location` = $\mu$ and `scale` = $\lambda$ has density

$$
f(x \mid \mu, \lambda) = \frac{1}{2\lambda} exp\left(-\frac{\mid x - \mu \mid}{\lambda}\right)
$$

The mean and variance are

$$
E(X) = \mu; \quad Var(X) = 2 \lambda^2
$$
```{r, warning=FALSE, results='hide'}
library(rmutil) # for double-exponential 
library(LaplacesDemon)
```

Plotting the $Laplace(\mu, \lambda)$ distribution

```{r}
library(shiny)
library(ggplot2)
library(tidyverse)
# library(rmutil) # for double-exponential 
# library(LaplacesDemon)

ui <- fluidPage(
  column(4,
    wellPanel(
    sliderInput('mu','mu (location)', min = -3, max = 3, value = 0, step = 1),
    sliderInput('lambda','lambda (scale)', min = 0, max = 10, value = 1, step = 1)
    )
  ),
  column(8,
    plotOutput('plotDensity')
  )
)

server <- function(input,output) {

  mean <- reactive({input$mu})
  sd <- reactive({sqrt(2*input$lambda^2)})
  q9999 <- reactive({qlaplace(p = 0.9999, location = input$mu, scale = input$lambda)})

  selectedData <- reactive({
    data_frame(x = c(c(-q9999(),q9999())))
  })

  output$plotDensity <- renderPlot({
    ggplot(selectedData(), aes(x=x)) +
      geom_vline(xintercept = mean(), linetype = "dashed", alpha = 0.4) +
      stat_function(fun = dlaplace, args = c(location = input$mu, scale = input$lambda),
                    color = "#84CA72", size = 1) +
      labs(caption = paste0("Laplace (mu = ", input$mu, ", lambda = ", input$lambda,") Distribution")) +
      xlab("x") +
      ylab(paste0("dgamma(x, Laplace (mu = ", input$mu, ", lambda = ", input$lambda,")")) +
      coord_cartesian(ylim = c(0,0.5)) +
      theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      )

  })

}

shinyApp(ui = ui, server = server)
```


```{r, warning=FALSE, fig.align='center', results='hide', fig.show='hide'}
# dlaplace(x, location=0, scale=1, log=FALSE)
mu <- 0 # location
lambda <- 1/0.1 # scale

mean <- mu
sd <- sqrt(2*lambda^2)
q9999 <- qlaplace(p = 0.9999, location = mu, scale = lambda)

p <- data_frame(x = c(-q9999,q9999)) %>% 
  ggplot(aes(x=x))

p + geom_vline(xintercept = mean, linetype = "dashed", alpha = 0.4) +
    stat_function(fun = dlaplace, args = c(location = mu, scale = lambda),
                  color = "#84CA72", size = 1) +
    ggtitle(paste0("Laplace (mu = ", mu, ", lambda = ", lambda,") Distribution")) +
    xlab("x") +
    ylab(paste0("dlaplace(x, Laplace (mu = ", mu, ", lambda = ", lambda,")"))
```


```{r, warning=FALSE, fig.align='center'}
# rlaplace(n, location=0, scale=1, log=FALSE)
set.seed(1107)
n = 10000

mu <- 0 # location
lambda1 <- 1/0.1 # scale
ilambda1 <- 1/lambda1 # rate

tausq1 <-  rexp(n, rate = ilambda1^2/2)
mean(tausq1)


library(tictoc)
tic()
x1 <- rep(NA,n)
for (i in 1:n) {
  x1[i] <- rnorm(1, mean = mu, sd = sqrt(tausq1[i]))
}
toc()
c(mean(x1),sd(x1))


library(purrr)
tic()
x1 <- map_dbl(tausq1, ~ rnorm(1, mu, sd = sqrt(.x)))
toc()
c(mean(x1),sd(x1))


ilambda0 <- rgamma(n, shape = 1, rate = 0.00001)
c(mean(ilambda0),sd(ilambda0))

tausq0 <- rep(NA,n)
x0 <- rep(NA,n)
for (i in 1:n) {
  tausq0[i] <-  rexp(1, rate = ilambda0[i]^2/2)
  x0[i] <- rnorm(1, mean = mu, sd = sqrt(tausq0[i]))
}

mean(tausq0)
c(mean(x0), sd(x0))


p <- ggplot(data_frame(x1), aes(x=x1)) 
p + geom_density(color = 'red') +
  coord_cartesian(xlim = c(-100,100))

p <- ggplot(data_frame(x0), aes(x=x0)) 
p + geom_density(color = 'blue') +
  coord_cartesian(xlim = c(-0.05,0.05))


gamma <- rbinom(n, 1, prob = 0.5)
x <- gamma*x1 + (1-gamma)*x0
c(mean(x),sd(x))

p <- ggplot(data_frame(x), aes(x=x)) 
p + geom_density(color = 'green') 
```


```{r, warning=FALSE, fig.align='center'}
# rlaplace(n, location=0, scale=1, log=FALSE)
set.seed(1107)
n = 10000

ilambda <- rgamma(n, shape = 1, rate = 1)
mean(ilambda)
sd(ilambda)

tausq <-  rexp(n, rate = ilambda^2/2)
x <- rnorm(n, mean = mu, sd = sqrt(tausq))
mean(tausq)
sd(x)
mean(x)

p <- ggplot(data_frame(x), aes(x=x)) 
p + geom_density(color = 'green') 
```

# Bayesian Shrinkage Priors
## `CTS-studt`

```{r, warning=FALSE, fig.align='center'}

# Student-t distribution ====

# df <- 6
# ncp <- NA # non-centrality parameter
# 
# mean <- 0
# sd <- sqrt(df/(df-2))
# 
# p <- data_frame(x = c(mean - 4*sd, mean + 4*sd)) %>%
#   ggplot(aes(x=x))
# 
# p + geom_vline(xintercept = mean, linetype = "dashed", alpha = 0.4) +
#     stat_function(fun = dt, args = c(df),
#                   color = "#84CA72", size = 1) +
#     labs(caption = paste0("t Distribution (df = ", df,")")) +
#     xlab("x") +
#     ylab(paste0("dt(x, df = ", df,")")) +
#     coord_cartesian(ylim = c(0,0.5)) +
#     theme(
#         plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
#       )

dt2 <- function(df) {
  function (x) dt(x, df = df)
}

stat_studt <- function(df) {
  stat_function(aes(colour = df), fun = dt2(df), size = 0.5)
}

ggplot(data_frame(x = c(-5,5)), aes(x)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.4) +
  lapply(c(2,4,6), stat_studt) +
  scale_colour_viridis_c(limits = c(2,6)) +
    labs(caption = paste0("t Distribution (df)")) +
    xlab("x") +
    ylab(paste0("dt(x, df)")) +
    theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      )

# CTS-studt prior ====
ndraws <- 1e5

alpha <- 1
beta <- 0.001
tausq <- 1/rgamma(ndraws, shape = alpha, scale = 1/beta)
library(purrr)
x <- map_dbl(tausq, ~rnorm(1, mean = 0, sd = sqrt(.x)))


ggplot(data_frame(x), aes(x=x)) +
  geom_density(color = "red", fill = "lightgreen") +
  xlim(-10,10) + coord_cartesian(c(-5,5))+
    labs(caption = paste0("CTS-studt prior (alpha = ", alpha, ", beta = ", beta, ")")) +
    xlab("x") +
    ylab(paste0("density")) +
    theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      )

# ggplot(data_frame(x), aes(x=x)) +
#   stat_density(geom = "line", position = "identity") +
#   xlim(-10,10) + coord_cartesian(c(-5,5))

# Choose hyper-parameters
library(purrr)
ndraws <- 1e5

CTS_studt_pa1 <- function(alpha){
  set.seed(123)
  ndraws <- 1e5
  beta <- 0.001
  tausq <- 1/rgamma(ndraws, shape = alpha, scale = 1/beta)
  x <- map_dbl(tausq, ~rnorm(1, mean = 0, sd = sqrt(.x)))
  return(x)
}

df1_pa1 <- lapply(c(1,2,3), CTS_studt_pa1)
df2_pa1 <- data_frame(x = unlist(df1_pa1), alpha = factor(rep(c(1,2,3), each = ndraws)))

plot_pa1 <- ggplot(df2_pa1, aes(x=x, color = alpha)) +
  geom_density() +
  xlim(-10,10) + coord_cartesian(c(-5,5))

tail_pa1 <- ggplot(df2_pa1, aes(x=x, color = alpha)) +
  geom_line(aes(y = 1 - ..y.., linetype = alpha), stat='ecdf') +
  scale_linetype_manual(values = c(1, 2, 3)) +
  coord_cartesian(xlim = c(0, 1.5), ylim = c(0, 0.5), expand = FALSE) +
  theme(axis.title = element_blank(), legend.position = "bottom") + 
  guides(lty=guide_legend(nrow=1,byrow=TRUE))

CTS_studt_pa2 <- function(beta){
  set.seed(123)
  ndraws <- 1e5
  alpha <- 1
  tausq <- 1/rgamma(ndraws, shape = alpha, scale = 1/beta)
  x <- map_dbl(tausq, ~rnorm(1, mean = 0, sd = sqrt(.x)))
  return(x)
}

df1_pa2 <- lapply(c(0.001,0.01,0.1), CTS_studt_pa2)
df2_pa2 <- data_frame(x = unlist(df1_pa2), beta = factor(rep(c(0.001,0.01,0.1), each = ndraws)))

plot_pa2 <- ggplot(df2_pa2, aes(x=x, color = beta)) +
  geom_density() +
  xlim(-10,10) + coord_cartesian(c(-5,5))

tail_pa2 <- ggplot(df2_pa2, aes(x=x, color = beta)) +
  geom_line(aes(y = 1 - ..y.., linetype = beta), stat='ecdf') +
  scale_linetype_manual(values = c(1, 2, 3)) +
  coord_cartesian(xlim = c(0, 1.5), ylim = c(0, 0.5), expand = FALSE) +
  theme(axis.title = element_blank(), legend.position = "bottom") + 
  guides(lty=guide_legend(nrow=1,byrow=TRUE))

# xlim removes values outside the range, coord_cartesian does not, but behaves strange sometimes
# solution: restrict range with xlim first & then use coord-cartesian  
# percentage of obs removed
sum((df2_pa2$x < 10) & (df2_pa2$x > -10))/length(df2_pa2$x)*100


library(gridExtra)
grid.arrange(plot_pa1,plot_pa2, ncol = 2)
# tail behavior
grid.arrange(tail_pa1,tail_pa2, ncol = 2)
```

## `CTS-lasso`

```{r, warning=FALSE, fig.align='center'}
# Laplace distribution ====
library(LaplacesDemon)

# rlaplace(n, location=0, scale=1, log=FALSE)

# mu <- 0 # location
# lambda <- 1/0.1 # scale
# 
# x <- rlaplace(n = 1e5, location = mu, scale = lambda)
# 
# ggplot(data_frame(x), aes(x=x)) +
#   geom_density(fill = "lightgreen") +
#   stat_function(fun = dlaplace, args = c(location = mu, scale = lambda),
#                   color = "red", size = 1) 



dlaplace2 <- function(scale) {
  function (x) dlaplace(x, location = 0, scale = scale)
}

stat_laplace <- function(scale) {
  stat_function(aes(colour = scale), 
                fun = dlaplace2(scale = scale), 
                size = 0.5) 
}

ggplot(data_frame(x = c(-5,5)), aes(x=x)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.4) +
  lapply(c(1,5,10), stat_laplace) +
  scale_colour_viridis_c(limits = c(1,10)) +
    labs(caption = paste0("Laplace Distribution (scale)")) +
    xlab("x") +
    ylab(paste0("dlapalce(x, scale)")) +
    theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      )


# CTS-lasso prior ====

# Double exponential
ndraws <- 1e5
mu <- 0 # location
scale <- 1/1 # scale
lambdasq <- 1/scale # rate

tausq <-  rexp(ndraws, rate = lambdasq/2)
library(purrr)
x <- map_dbl(tausq, ~ rnorm(1, mean = mu, sd = sqrt(.x)))

ggplot(data_frame(x), aes(x=x)) +
  geom_density(fill = "lightgreen") +
  stat_function(fun = dlaplace, args = c(location = mu, scale = scale),
                  color = "red", size = 1)

# Hierarchical 
ndraws <- 1e5

library(purrr)
alpha <- 1
sigma <- 1

lambdasq <- rgamma(ndraws, shape = alpha, scale = sigma)
tausq <-  map_dbl(lambdasq, ~rexp(1, rate = .x/2))
x <- map_dbl(tausq, ~ rnorm(1, mean = 0, sd = sqrt(.x)))

# c(mean(lambdasq), sd(lambdasq))
scale <- 1/sqrt(mean(lambdasq)) # for corresponding Laplace distribution

ggplot(data_frame(x), aes(x=x)) +
  geom_density(fill = "lightgreen") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.4) +
  stat_function(fun = dlaplace, args = c(location = 0, scale = scale),
                  color = "red", size = 0.5) +
  xlim(-10,10) + coord_cartesian(c(-5,5)) +
  labs(caption = paste0("CTS-lasso prior (alpha = ", alpha, ", sigma = ", sigma, ")")) +
  xlab("x") +
  ylab(paste0("density")) +
  theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
  )


# Choose hyper-parameters
library(purrr)
ndraws <- 1e5

CTS_lasso_pa1 <- function(alpha) {
  set.seed(123)
  sigma = 1
  lambdasq <- rgamma(ndraws, shape = alpha, scale = sigma)
  tausq <-  map_dbl(lambdasq, ~rexp(1, rate = .x/2))
  x <- map_dbl(tausq, ~ rnorm(1, mean = 0, sd = sqrt(.x)))
  return(x)
}

df1_pa1 <- lapply(c(1,2,3), CTS_lasso_pa1)
df2_pa1 <- data_frame(x = unlist(df1_pa1), alpha = factor(rep(c(1,2,3), each = ndraws)))

plot_pa1 <- ggplot(df2_pa1, aes(x=x, color = alpha)) +
  geom_density() +
  xlim(-10,10) + coord_cartesian(c(-5,5))

tail_pa1 <- ggplot(df2_pa1, aes(x=x, color = alpha)) +
  geom_line(aes(y = 1 - ..y.., linetype = alpha), stat='ecdf') +
  scale_linetype_manual(values = c(1, 2, 3)) +
  coord_cartesian(xlim = c(0, 2), ylim = c(0, 0.5), expand = FALSE) +
  theme(axis.title = element_blank(), legend.position = "bottom") + 
  guides(lty=guide_legend(nrow=1,byrow=TRUE))

CTS_lasso_pa2 <- function(sigma) {
  set.seed(123)
  alpha = 1
  lambdasq <- rgamma(ndraws, shape = alpha, scale = sigma)
  tausq <-  map_dbl(lambdasq, ~rexp(1, rate = .x/2))
  x <- map_dbl(tausq, ~ rnorm(1, mean = 0, sd = sqrt(.x)))
  return(x)
}

df1_pa2 <- lapply(c(1,2,3), CTS_lasso_pa2)
df2_pa2 <- data_frame(x = unlist(df1_pa2), sigma = factor(rep(c(1,2,3), each = ndraws)))

plot_pa2 <- ggplot(df2_pa2, aes(x=x, color = sigma)) +
  geom_density() +
  xlim(-10,10) + coord_cartesian(c(-5,5))

tail_pa2 <- ggplot(df2_pa2, aes(x=x, color = sigma)) +
  geom_line(aes(y = 1 - ..y.., linetype = sigma), stat='ecdf') +
  scale_linetype_manual(values = c(1, 2, 3)) +
  coord_cartesian(xlim = c(0, 2), ylim = c(0, 0.5), expand = FALSE) +
  theme(axis.title = element_blank(), legend.position = "bottom") + 
  guides(lty=guide_legend(nrow=1,byrow=TRUE))

# xlim removes values outside the range, coord_cartesian does not, but behaves strange sometimes
# solution: restrict range with xlim first & then use coord-cartesian  
# percentage of obs removed
# sum((df2_pa1$x < 10) & (df2_pa1$x > -10))/length(df2_pa1$x)*100
# sum((df2_pa2$x < 10) & (df2_pa2$x > -10))/length(df2_pa2$x)*100

library(gridExtra)
grid.arrange(plot_pa1,plot_pa2, ncol = 2)
# tail behavior
grid.arrange(tail_pa1,tail_pa2, ncol = 2)
```

## `CTS-horse-mx`

```{r}
ndraws = 1e5
library(purrr)

nu <- 1/rgamma(ndraws, shape = 1/2, scale = 1)
lamdasq <- map_dbl(nu, ~ 1/rgamma(1, shape = 1/2, scale = .x))
xi <- 1/rgamma(ndraws, shape = 1/2, scale = 1)
tausq <- map_dbl(xi, ~ 1/rgamma(1, shape = 1/2, scale = .x))
temp <- lamdasq*tausq
x <- map_dbl(temp, ~rnorm(1, mean = 0, sd = .x))

ggplot(data_frame(x), aes(x=x)) +
  geom_density(color = "red", fill = "lightgreen") +
    xlim(-10,10) + coord_cartesian(c(-5,5))+
    labs(caption = paste0("CTS-horseshoe-mx prior")) +
    xlab("x") +
    ylab(paste0("density")) +
    theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      )
```


## `CTS-horse-sl`

```{r}
ndraws = 1e5
library(purrr)
lambda <- rhalfcauchy(ndraws, scale= 1)
tau <- map_dbl(lambda, ~rhalfcauchy(1, scale = .x))
x <- map_dbl(tau, ~rnorm(1, mean = 0, sd = .x))

ggplot(data_frame(x), aes(x=x)) +
  geom_density(color = "red", fill = "lightgreen") +
    xlim(-10,10) + coord_cartesian(c(-5,5))+
    labs(caption = paste0("CTS-horseshoe-sl prior")) +
    xlab("x") +
    ylab(paste0("density")) +
    theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      )
```



## `SSVS-normal`

## `SSVS-studt`
```{r}
ndraws <- 1e5

# SSVS-studt1
alpha1 <- 1
beta1 <- 1
tausq1 <- 1/rgamma(ndraws, shape = alpha1, scale = 1/beta1)
library(purrr)
studt_slab <- map_dbl(tausq1, ~rnorm(1, mean = 0, sd = sqrt(.x)))

tausq0 <- 1e-3
library(purrr)
studt_spike <- map_dbl(tausq0, ~rnorm(1, mean = 0, sd = sqrt(.x)))


# SSVS-studt2
alpha1 <- 1
beta1 <- 1
tausq1 <- 1/rgamma(ndraws, shape = alpha1, scale = 1/beta1)
library(purrr)
studt_slab <- map_dbl(tausq1, ~rnorm(1, mean = 0, sd = sqrt(.x)))

tausq0 <- tausq1*1e-3
library(purrr)
studt_spike <- map_dbl(tausq0, ~rnorm(1, mean = 0, sd = sqrt(.x)))
#mean(tausq0) #0.012


# SSVS-studt3
alpha1 <- 1
beta1 <- 1
tausq1 <- 1/rgamma(ndraws, shape = alpha1, scale = 1/beta1)
library(purrr)
studt_slab <- map_dbl(tausq1, ~rnorm(1, mean = 0, sd = sqrt(.x)))

alpha0 <- 1
beta0 <- 0.0001
tausq0 <- 1/rgamma(ndraws, shape = alpha0, scale = 1/beta0)
library(purrr)
studt_spike <- map_dbl(tausq0, ~rnorm(1, mean = 0, sd = sqrt(.x)))
#mean(tausq0) #0.0035


pi <- rbern(1e5, prob = 0.2)
x <- pi*studt_slab + (1-pi)*studt_spike
# tmp <- runif(1e5)
# x <- (tmp<=0.2)*studt_slab + (tmp>0.2)*studt_spike
ggplot(data_frame(x), aes(x=x)) +
  geom_density(color = "red", fill = "lightgreen") +
  xlim(-10,10) + coord_cartesian(c(-5,5))+
    labs(caption = paste0("SSVS-studt prior")) +
    xlab("x") +
    ylab(paste0("density")) +
    theme(
        plot.caption = element_text(hjust = 0.5, size = 14, face ="italic")
      )
```

## `SSVS-lasso`

## `SSVS-horse`

```{r}
##### Sample from prior densities -----

set.seed(123)
ndraws <- 1e5

## ridge 
normal <- rnorm(ndraws, mean=0, sd=0.5)

## Student's t 
studt <- rt(ndraws, df=1) 

CTS_studt_fun <- function(alpha = 1, beta = 0.001, ndraws = 1e5) {
  library(purrr)
  tausq <- 1/rgamma(ndraws, shape = alpha, scale = 1/beta)
  x <- map_dbl(tausq, ~rnorm(1, mean = 0, sd = sqrt(.x)))
  return(x)
}
CTS_studt <- CTS_studt_fun(alpha = 1, beta = 0.001, ndraws = 1e5)


## lasso 
lasso <- rmutil::rlaplace(ndraws, m=0, s=0.5)

CTS_lasso_fun <- function (alpha = 1, sigma = 1, ndraws = 1e5) {
  library(purrr)
  lambdasq <- rgamma(ndraws, shape = alpha, scale = sigma)
  tausq <-  map_dbl(lambdasq, ~rexp(1, rate = .x/2))
  x <- map_dbl(tausq, ~ rnorm(1, mean = 0, sd = sqrt(.x)))
  return(x)
}
CTS_lasso <- CTS_lasso_fun(alpha = 1, sigma = 1, ndraws = 1e5)

## horseshoe

CTS_horsemx_fun <- function(ndraws = 1e5) {
  library(purrr)
  nu <- 1/rgamma(ndraws, shape = 1/2, scale = 1)
  lamdasq <- map_dbl(nu, ~ 1/rgamma(1, shape = 1/2, scale = .x))
  xi <- 1/rgamma(ndraws, shape = 1/2, scale = 1)
  tausq <- map_dbl(xi, ~ 1/rgamma(1, shape = 1/2, scale = .x))
  temp <- lamdasq*tausq
  x <- map_dbl(temp, ~rnorm(1, mean = 0, sd = .x))
  return(x)
}
CTS_horsemx <- CTS_horsemx_fun(ndraws = 1e5)

CTS_horsesl_fun <- function(ndraws = 1e5) {
  library(purrr)
  lambda <- rhalfcauchy(ndraws, scale= 1)
  tau <- map_dbl(lambda, ~rhalfcauchy(1, scale = .x))
  x <- map_dbl(tau, ~rnorm(1, mean = 0, sd = .x))
  return(x)
}
CTS_horsesl <- CTS_horsesl_fun(ndraws = 1e5)
```


```{r}
##### Plot -----
## Create plot data
df_comb <- data_frame(CTS_studt, CTS_lasso, CTS_horsemx, CTS_horsesl)
df_long <- gather(df_comb, Prior, value) # long format
df_long$Prior <- factor(df_long$Prior)
levels(df_long$Prior) <- list("CTS_studt" = "CTS_studt", "CTS_lasso" = "CTS_lasso", "CTS_horsemx" = "CTS_horsemx", "CTS_horsesl" = "CTS_horsesl")
df_long$asymp <- rep(NA, nrow(df_long)) 
df_long[which(df_long$Prior=="CTS_horsemx"|df_long$Prior=="CTS_horsesl"), "asymp"] <- 0

   
df_long <- gather(df_comb, Prior, value) %>% 
      mutate(Prior = factor(Prior, levels = list("CTS_studt" = "CTS_studt", "CTS_lasso" = "CTS_lasso", "CTS_horsemx" = "CTS_horsemx", "CTS_horsesl" = "CTS_horsesl")),
             asymp = case_when(Prior=="CTS_horsemx" ~ 0,
                               Prior=="CTS_horsesl" ~ 0,
                               TRUE ~ NA_real_))

plotComb <- ggplot(df_long, aes(x = value, colour = Prior, linetype = Prior)) +
  stat_density(geom = "line", position = "identity") +
  geom_vline(aes(xintercept = asymp), colour = "grey") +
  scale_linetype_manual(values = c(1,2,3,4)) +
  scale_color_manual(values = c(1,2,3,4)) +
  xlim(-10,10) + coord_cartesian(xlim=c(-5,5), ylim =c(0,5)) +
  theme(axis.title = element_blank(), legend.position = "bottom")
# xlim removes values outside the range, coord_cartesian does not, but behaves strange sometimes
# solution: restrict range with xlim first & then use coord-cartesian  

##### Tail behavior -----
# Plot the survival function, i.e., 1-CDF
# See https://stats.stackexchange.com/questions/86429/which-has-the-heavier-tail-lognormal-or-gamma

tailComb <- ggplot(df_long, aes(x = value, group = Prior)) +
  geom_line(aes(y = 1 - ..y.., linetype = Prior, colour = Prior), stat='ecdf') +
  scale_linetype_manual(values = c(1,2,3,4)) +
  scale_color_manual(values = c(1,2,3,4)) +
  coord_cartesian(xlim = c(0, 8), ylim = c(0, 0.5), expand = FALSE) +
  theme(axis.title = element_blank(), legend.position = "bottom", legend.title = element_blank()) + 
  guides(lty=guide_legend(nrow=1,byrow=TRUE))

library(gridExtra)
get_legend <- function(myggplot){
  require(gridExtra)
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend <- get_legend(plotComb)
plotComb <- plotComb + theme(legend.position="none")
tailComb <- tailComb + theme(legend.position="none")
grid.arrange(plotComb,tailComb, legend, ncol = 2, nrow =2,
             layout_matrix = rbind(c(1,2),c(3,3)),
             heights = c(4,1))
```






